Automatically generating alternative hypotheses and controls

### 2018-02-22 topics.tex
topics Pareto frontier writing

@@

A successful model of artificial grammar learning should also explain
at least some aspects of first language acquisition.  As a first step
in this direction, we take the same model that we applied to these AGL
learning problems and use it to solve phonology textbook exercises.
Each of these textbook exercises is an inductive reasoning problem
where the learner must explain surface pronunciations in terms of a
latent generative model (a grammar).
Figure~\ref{naturalLanguageExample} shows our model solving a
representative textbook problem.  We believe this extension two
natural language is an important next step for models of artificial
grammar learning,
and helps deliver on the basic promises and presuppositions of AGL research.

>>>>

As an example of how the Pareto front offers a lens on the
space of possible hypotheses,
consider the upper left Pareto front in figure~\ref{quadrupleFrontier} (\emph{aab, 1 example}).
Here the learner has seen a single word, [\textipa{vEfefe}].
Some grammars living on the Pareto frontier are:
\begin{itemize}
\item \textbf{A grammar that does nothing:} In the lower right-hand corner of~\ref{quadrupleFrontier}
  is the grammar labeled ``surface=underlying'',
  which just says that every word (a ``surface pronunciation'')
  is represented (``underlyingly'') literally how it is pronounced.
  The observed word [\textipa{vEfefe}] is represented as /\textipa{vEfefe}/, which has 6 symbols,
  giving a fit to the data of -6.
\item \textbf{A grammar that duplicates syllables:} Highlighted in dark red in~\ref{quadrupleFrontier}
  is a grammar with the single rule $\varnothing\to\sigma_i/\_\sigma_i$\texttt{\#},
  which inserts a copy of the last syllable syllable.
  This rule generates the word [\textipa{vEfefe}] by starting with the stem (i.e., underlying form) /\textipa{vEfe}/ (which has 4 symbols, fit to data of -4) and then applying the rule in the grammar,
  which copies the last syllable and makes [\textipa{vEfefe}].
\item \textbf{A grammar that duplicates syllables and appends
  morphemes:} Highlighted in pink in~\ref{quadrupleFrontier} are
  grammars that duplicate a syllable but also append or prepend extra
  morphemes.  These correspond to generalizations where the learner
  believes that different parts of [\textipa{vEfefe}] correspond to
  prefixes or suffixes in the language.  For example, the grammar in
  the upper left corner incorporates the suffix /\textipa{fe}/ as well
  as the prefix /\textipa{vE}/, and explains the observation using an
  underlying form that is completely empty (fit to data of 0) -- this
  grammar has memorized the observed data, and so maximally compresses
  it, but at the cost of having many symbols inside of the grammar (22 symbols, vs 6 symbols in the grammar that just duplicates a syllable).  
  \end{itemize}

@@

A successful model of artificial grammar learning should also explain
at least some aspects of first language acquisition.  As a first step
in this direction, we take the same model that we applied to these AGL
learning problems and use it to solve phonology textbook exercises.
Each of these textbook exercises is an inductive reasoning problem
where the learner must explain surface pronunciations in terms of a
latent generative model (a grammar).

These natural language results are still pulmonary and
ongoing. Figure~\ref{naturalLanguageExample} shows our model solving a
representative textbook problem.  We believe this extension two
natural language is an important next step for models of artificial
grammar learning, and helps deliver on the basic promises and
presuppositions of AGL research.
